An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-10-12 aistats2017.tex
Shortened paper more, abotu 8.4 pages. We had repetitive info in sections 1/2

    \item Increase the minibatch size when needed for certain minibatches. Log
    acceptance variance scales as $1/k$ vs the minibatch size $k$. Our test is
    adaptive like earlier works, but unlike them, the distribution of minibatch
    size is Gaussian, not long-tailed.  Increased minibatch size also reduces
    the error rate for the test.

    \item For continuous probability distributions, reduce the proposal step
    size and variance (for stochastic proposals) compared to an optimal
    proposal. The variance of the log acceptance probability scales as the
    square of proposal step size. 
\end{itemize}

It is worth discussing at this point the typical goals of M-H sampling on large
datasets. By the Bernstein-von Mises Theorem, the posterior distribution for a
Bayesian inference task has variance that scales inversely with the $N$ data
samples. Simply sampling from it is one application, but an efficient
proposal~\citep{OptimalScaling01} has similar variance to the target and will
diffuse to it extremely slowly. For applications to neural networks or models
where the posterior is multimodal~\citep{choromanska2014loss}, samplers will
likely get trapped in one of the modes. A common solution is to anneal the
sampler, running first at high temperature (scaling log likelihoods by $1/T$)
to flatten the likelihood landscape. This in turn reduces the variance of
the log acceptance probability and allows our acceptance test to be applied. Our
samples can cover the search space densely with small steps rather than taking a
few sparse steps towards an optimum. In this mode, Metropolis-Hastings can be
used in similar fashion to Stochastic Gradient Descent. The goal in SGD is to
make gradual progress to a posterior mode with each step, taking small steps so
that the cumulative displacement has progressively lower variance.

% Daniel: I merged these two paragraphs into one, because these were too long,
% and we also never experiment with the step size. Let me know what you think.
\begin{comment}
It is worth discussing at this point the typical goals of M-H sampling on large
datasets.  By the Bernstein-von Mises Theorem, the posterior distribution for a
Bayesian inference task is asymptotically normal, and has variance that scales
inversely with the $N$ data samples. Simply sampling from this distribution is
one application, but an efficient proposal~\citep{OptimalScaling01} has similar
variance to the target distribution and will diffuse to it extremely slowly from
an initialization value which is (likely to be) many standard deviations away.
If there are any other strong modes, it is very likely for the sampler to find
one of them and become trapped in it when run at the normal distribution
temperature ($T=1$). A common solution is to anneal the sampler, running first
at high temperature (scaling log likelihoods by $1/T$) which flattens the
likelihood landscape.  This in turn reduces the variance of the log acceptance
probability and allows our acceptance test to be applied.

A second question concerns step size. Once we have fixed temperature, our
variance constraint implies that we have to trade-off proposal step size $s$ and
batch size $b$ ($b \propto p^2$), i.e. we can make many small steps, or one
large step, with a given batch of data. One of the primary drivers of this work
is our belief in the value of small steps. For applications to neural networks
or other models where the posterior is multimodal, posterior inference is
arguably a search process. Covering the search space densely with small steps is
much more valuable than few sparse steps toward the nearest optimum. In this
mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
Descent. The goal in SGD is to make gradual progress to a posterior mode with
each step, taking small steps so that the cumulative displacement has
progressively lower variance. We demonstrate this behavior in our logistic
regression experiments. 
\end{comment}

@@

The acceptance test guarantees detailed balance, which means
$p(\theta)p(\theta'\mid\theta) = p(\theta')p(\theta \mid\theta')$, where
$p(\theta'\mid\theta)$ is the probability of a transition from state $\theta$ to
$\theta'$. Here, $p(\theta'\mid\theta)=q(\theta'\mid\theta)\alpha(\theta,\theta')$.
% Daniel: I commented this out because people should know what it means and we
%re-state another detailed balance equation later!
%\begin{equation}\label{eq:detailed_balance2}
%    p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
%\end{equation}
This condition, together with ergodicity, guarantees that the Markov chain has a
unique stationary distribution $\pi(\theta) = p(\theta)$. For Bayesian
inference, the target distribution is $p(\theta \mid x_1, \ldots, x_N)$. The
acceptance probability is now:
\begin{equation}\label{eq:acceptance_probability}
    \alpha(\theta,\theta') = 
    \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)}
    \wedge 1
\end{equation}
where $p_0(\theta)$ is the prior. Computing samples this way requires the use of all $N$
training data points, but this is very expensive for large datasets. To address
this challenge,~\citet{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
Metropolis-Hasting tests using sequential hypothesis testing. During each
iteration, they start with a small minibatch of data and test whether the sample $\theta'$ 
should be accepted based on an approximate version of
the test $u < \alpha(\theta,\theta')$. If the approximate test cannot make a
decision with sufficient confidence, then the minibatch size is increased and
the test repeats. This process continues until a decision. The bounds depend on
either an asymptotic Central Limit Theorem~\citep{cutting_mh_2014} or a
concentration bound~\citep{icml2014c1_bardenet14}. The latter requires direct
bounds on the log likelihood ratio, which for general distributions requires
knowing $p(x_i \mid \theta)$ and $p(x_i \mid \theta')$ for all $N$ samples. In
addition, both methods suffer the drawback of resolving small log likelihood
ratio differences between the minibatch and full batch versions.  We discuss
a worst-case scenario in Section~\ref{ssec:gaussian_example}.

@@

Since minibatches contains randomly selected samples $x_i$, the values
$\Lambda_i$ are i.i.d. random
variables\footnote{The analysis assumes sampling with replacement
although implementations on typical large datasets will approximate
this by sampling without replacement.}.
By the Central Limit Theorem, we expect $\Lambda^*(\theta,\theta')$ to
be approximately Gaussian. The acceptance test then becomes a
statistical test of the hypothesis that
$\Lambda(\theta,\theta')>\psi(u,\theta,\theta')$ by establishing that
$\Lambda^*(\theta,\theta')$ is substantially larger than
$\psi(u,\theta,\theta')$.  In~\citet{cutting_mh_2014} an asymptotic
central limit argument was used to derive this gap, while
in~\citet{icml2014c1_bardenet14} a concentration bound was used. In
both cases, the resulting tests were shown to give useful reductions
in number of samples required over using the full dataset, but there
were no worst-case bounds, which we explore in the following section.

@@

\begin{lemma}\label{lem:worst_case}
    For the model in Section~\ref{ssec:gaussian_example}, there exists a fixed
    (independent of $N$) constant $c$ such that with probability $\geq c$ over
    the joint distribution of $(\theta, \theta', u)$, the tests
    from~\cite{cutting_mh_2014,icml2014c1_bardenet14} consume all $N$ samples. 
\end{lemma}
\vspace{-1em}
\begin{proof}
See Supplementary Material, Section~\ref{app:worst_case_proof}.
\end{proof}
Similar results can be shown for other distributions and proposals by
identifying regions in product space $(\theta,\theta',u)$ such that the
hypothesis test needs to separate nearly-equal values.  It follows that the
accelerated M-H tests from prior work require at least a constant fraction $\geq
c$ in the amount of data consumed per test compared to full-data tests, so their
speed-up is $\le 1/c$.

The issue is the use of tail bounds to separate $\Delta$ from zero; for certain
input/random $u$ combinations, $\Delta$ can be arbitrarily close to zero. We
avoid this by using the {\em approximately normal} variation in $\Delta^*$ to
{\em replace} the variation due to $u$.

@@

\subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\citep{Barker65} function:
$g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it satisfies the
condition in Lemma~\ref{lem:detailed_balance}.  While it is slightly less
efficient than the Metropolis test when used on the full dataset, we will see
that its smoothness allows it to naturally tolerate substantial variance in its
input argument. This in turn will lead to a much more efficient test on subsets
of data.

@@

Our approach is to instead decompose $X_{\rm log}$ as
\begin{equation}\label{eq:deconvolution}
    X_{\rm log} = X_{\rm norm}+X_{\rm corr},
\end{equation}
where we assume $X_{\rm norm} \sim \mathcal{N}(0, \sigma^2)$ and that $X_{\rm
corr}$ is a zero-mean ``correction'' variable with density $C_{\sigma}(X)$.  The
two variables are added (i.e., their distributions convolve) to form $X_{\rm
log}$.  This decomposition requires an appropriate $X_{\rm corr}$ distribution.
We show in Section~\ref{sec:correction} that it is possible to use deconvolution
to derive a highly accurate representation of $C_{\sigma}(X)$.  Using $X_{\rm
corr}$ samples from $C_{\sigma}(X)$, the acceptance test is now
\begin{equation}\label{eq:criteria}
    \Delta + X_{\rm log} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
\end{equation}
Therefore, assuming the variance of $\Delta^*$ is small enough, if we have an
estimate of $\Delta^*$ from the current data minibatch, we test acceptance by
adding a random variable $X_{\rm corr}$ and then accept $\theta'$ if the result
is positive (and reject otherwise).

@@

Our test in Equation~(\ref{eq:criteria}) requires knowing the
distribution of $X_{\rm corr}$ such that $X_{\rm norm} +
X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \sim \mathcal{N}(0,\sigma^2)$
and $X_{\rm log}$ has a standard logistic CDF, $(1+\exp(-X))^{-1}$. In
Section~\ref{sec:analysis}, we show that the test accuracy depends on the
absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and $X_{\rm
log}$. Consequently, we need to minimize this in our construction of $X_{\rm
corr}$. More formally, let $\Phi_{s_X} = \Phi(X/s_X)$ where $\Phi$ is the
standard normal CDF\footnote{Hence, $\Phi_{s_X}$ is the CDF of a zero-mean
Gaussian with standard deviation $s_X$.}, $S(X)$ be the logistic function, and
$C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
distribution. Our goal is to solve:
\begin{equation}\label{eq:overall_corr_problem}
    C_\sigma^* = \argmin_{C_\sigma} |\Phi_{\sigma} * C_{\sigma} - S|
\end{equation}
where $*$ denotes convolution.

@@

With this approach, there is no guarantee that $u^* \geq 0$. However, we have
some flexibility in the choice of $\sigma$ in Equation~(\ref{eq:overall_corr_problem}).
As we decrease the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$
grows by the same amount and is in fact the result of convolution with a
Gaussian whose variance is the difference.  Thus as $\sigma$ decreases,
$C_\sigma(X)$ grows and approaches the derivative of a logistic function at
$\sigma = 0$. It retains some very weak negative values for $\sigma > 0$ but
removal of those values leads to very small error. Table~\ref{tab:xcorr} shows
that the errors between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ can be
made very small, approaching single floating precision (about $10^{-7}$).

Algorithm~\ref{alg:our_algorithm} describes our procedure. A few points:

@@

We run MCMC with our MH test and benchmark with~\citet{cutting_mh_2014}
and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.  For
the former, we increment minibatches by 100 and for the latter, we increase it
geometrically with a ratio of $\gamma = 1.5$.  The tolerance for making a
decision in~\citet{cutting_mh_2014} is $\epsilon=0.005$, and the error bound
control parameters in~\citet{icml2014c1_bardenet14} are $p = 2$ and $\delta =
0.01$.  All methods (including ours) use the same random walk proposer with
covariance $\Sigma={\rm diag}(0.15, 0.15)$.  Since the proposer generates a
random walk, all shaping of distribution of the samples is due to the MH test.
All methods are run 5000 times to collect 5000 samples.

@@

We first observe that
\[
    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
\]
and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that for all $z$:
\begin{align*}
-\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
\end{align*}
as desired.

>>>>

    \item Increase the minibatch size when needed. Log acceptance variance
    scales as $1/k$ vs the minibatch size $k$. Our test is adaptive like earlier
    works, but unlike them, the distribution of minibatch size is Gaussian, not
    long-tailed.  Increased minibatch size also reduces the error rate for the
    test.

    \item For continuous distributions, reduce the proposal step size and
    variance compared to an optimal proposal. The variance of the log acceptance
    probability scales as the square of proposal step size. 
\end{itemize}

It is worth discussing at this point the typical goals of M-H sampling on large
datasets. By the Bernstein-von Mises Theorem, the posterior distribution for a
Bayesian inference task has variance that scales inversely with $N$. Simply
sampling from it is one application, but an efficient
proposal~\citep{OptimalScaling01} has similar variance to the target and will
diffuse to it extremely slowly. For applications to neural networks or models
where the posterior is multimodal~\citep{choromanska2014loss}, samplers will
likely get trapped in one of the modes. A common solution is to anneal the
sampler, running first at high temperatures to flatten the likelihood landscape.
This in turn reduces the variance of the log acceptance probability and allows
our test to be applied. Our samples can cover the search space densely with
small steps rather than taking a few sparse steps towards an optimum. In this
mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
Descent. The goal in SGD is to make gradual progress to a posterior mode with
each step, taking small steps so that the cumulative displacement has
progressively lower variance.

% Daniel: I merged these two paragraphs into one, because these were too long,
% and we also never experiment with the step size. Let me know what you think.

%% It is worth discussing at this point the typical goals of M-H sampling on large
%% datasets.  By the Bernstein-von Mises Theorem, the posterior distribution for a
%% Bayesian inference task is asymptotically normal, and has variance that scales
%% inversely with the $N$ data samples. Simply sampling from this distribution is
%% one application, but an efficient proposal~\citep{OptimalScaling01} has similar
%% variance to the target distribution and will diffuse to it extremely slowly from
%% an initialization value which is (likely to be) many standard deviations away.
%% If there are any other strong modes, it is very likely for the sampler to find
%% one of them and become trapped in it when run at the normal distribution
%% temperature ($T=1$). A common solution is to anneal the sampler, running first
%% at high temperature (scaling log likelihoods by $1/T$) which flattens the
%% likelihood landscape.  This in turn reduces the variance of the log acceptance
%% probability and allows our acceptance test to be applied.
%% 
%% A second question concerns step size. Once we have fixed temperature, our
%% variance constraint implies that we have to trade-off proposal step size $s$ and
%% batch size $b$ ($b \propto p^2$), i.e. we can make many small steps, or one
%% large step, with a given batch of data. One of the primary drivers of this work
%% is our belief in the value of small steps. For applications to neural networks
%% or other models where the posterior is multimodal, posterior inference is
%% arguably a search process. Covering the search space densely with small steps is
%% much more valuable than few sparse steps toward the nearest optimum. In this
%% mode, Metropolis-Hastings can be used in similar fashion to Stochastic Gradient
%% Descent. The goal in SGD is to make gradual progress to a posterior mode with
%% each step, taking small steps so that the cumulative displacement has
%% progressively lower variance. We demonstrate this behavior in our logistic
%% regression experiments.

@@

The acceptance test guarantees detailed balance, which means
$p(\theta)p(\theta'\mid\theta) = p(\theta')p(\theta \mid\theta')$, where
$p(\theta'\mid\theta)$ is the probability of a transition from state $\theta$ to
$\theta'$. Here, $p(\theta'\mid\theta)=q(\theta'\mid\theta)\alpha(\theta,\theta')$.
% Daniel: I commented this out because people should know what it means and we
%re-state another detailed balance equation later!
%\begin{equation}\label{eq:detailed_balance2}
%    p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
%\end{equation}
This condition, together with ergodicity, guarantees that the Markov chain has a
unique stationary distribution $\pi(\theta) = p(\theta)$. For Bayesian
inference, the target distribution is $p(\theta \mid x_1, \ldots, x_N)$. The
acceptance probability is now:
\begin{equation}\label{eq:acceptance_probability}
    \alpha(\theta,\theta') = 
    \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)}
    \wedge 1
\end{equation}
where $p_0(\theta)$ is the prior. Computing samples this way requires all $N$
data points, but this is very expensive for large datasets. To address this
challenge,~\citet{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
Metropolis-Hasting tests using sequential hypothesis testing. Each iteration,
they start with a small minibatch and test whether $\theta'$ should be accepted
based on approximating $u<\alpha(\theta,\theta')$. If the approximate test
cannot decide with sufficient confidence, the minibatch size is increased and
the test repeats. This process continues until a decision. The bounds depend on
either an asymptotic Central Limit Theorem~\citep{cutting_mh_2014} or a
concentration bound~\citep{icml2014c1_bardenet14}. The latter requires direct
bounds on the log likelihood ratio, which for general distributions requires
knowing $p(x_i \mid \theta)$ and $p(x_i \mid \theta')$ for all $N$ samples. In
addition, while both methods show useful reductions in the number of samples
required, they suffer the drawback of resolving small log likelihood ratio
differences between the minibatch and full batch versions. We discuss a
worst-case scenario in Section~\ref{ssec:gaussian_example}.

@@

Since minibatches contains randomly selected samples $x_i$, the values
$\Lambda_i$ are i.i.d. random variables\footnote{The analysis assumes sampling
with replacement although implementations on typical large datasets will
approximate this by sampling without replacement.}.  By the Central Limit
Theorem, we expect $\Lambda^*(\theta,\theta')$ to be approximately Gaussian. The
acceptance test then becomes a statistical test of the hypothesis that
$\Lambda(\theta,\theta')>\psi(u,\theta,\theta')$ by establishing that
$\Lambda^*(\theta,\theta')$ is substantially larger than
$\psi(u,\theta,\theta')$.

@@

\begin{lemma}\label{lem:worst_case}
    For the model in Section~\ref{ssec:gaussian_example}, there exists a fixed
    (independent of $N$) constant $c$ such that with probability $\geq c$ over
    the joint distribution of $(\theta, \theta', u)$, the tests
    from~\cite{cutting_mh_2014,icml2014c1_bardenet14} consume all $N$ samples. 
\end{lemma}
\vspace{-1em}
\begin{proof}
See Supplementary Material, Section~\ref{app:worst_case_proof}.
\end{proof}
Similar results can be shown for other distributions and proposals by
identifying regions in product space $(\theta,\theta',u)$ such that the
hypothesis test needs to separate nearly-equal values.  It follows that the
accelerated tests from prior work require at least a constant fraction $\geq c$
in the amount of data consumed per test compared to full-data tests, so their
speed-up is $\le 1/c$. The issue is the use of tail bounds to separate $\Delta$
from zero; for certain input/random $u$ combinations, $\Delta$ can be
arbitrarily close to zero. We avoid this by using the {\em approximately normal}
variation in $\Delta^*$ to {\em replace} the variation due to $u$.

@@

\subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\citep{Barker65} function:
$g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it satisfies the
condition in Lemma~\ref{lem:detailed_balance}.  While it is slightly less
efficient than the Metropolis test when used on the full dataset, we will see
that its smoothness allows it to naturally tolerate substantial variance in its
input. This in turn will lead to a much more efficient test on subsets of data.

@@

Our approach is to instead decompose $X_{\rm log}$ as
\begin{equation}\label{eq:deconvolution}
    X_{\rm log} = X_{\rm norm}+X_{\rm corr},
\end{equation}
where we assume $X_{\rm norm} \sim \mathcal{N}(0, \sigma^2)$ and that $X_{\rm
corr}$ is a zero-mean ``correction'' variable with density $C_{\sigma}(X)$.  The
two variables are added (i.e., their distributions convolve) to form $X_{\rm
log}$.  This decomposition requires an appropriate $C_\sigma$, which we derive
in Section~\ref{sec:correction}. Using $X_{\rm corr}$ samples from
$C_{\sigma}(X)$, the acceptance test is now
\begin{equation}\label{eq:criteria}
    \Delta + X_{\rm log} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
\end{equation}
Therefore, assuming the variance of $\Delta^*$ is small enough, if we have an
estimate of $\Delta^*$ from the current data minibatch, we test acceptance by
adding a random variable $X_{\rm corr}$ and then accept $\theta'$ if the result
is positive (and reject otherwise).

@@

Our test in Equation~(\ref{eq:criteria}) requires knowing the distribution of
$X_{\rm corr}$. In Section~\ref{sec:analysis}, we show that the test accuracy
depends on the absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$
and $X_{\rm log}$. Consequently, we need to minimize this in our construction of
$X_{\rm corr}$. More formally, let $\Phi_{s_X} = \Phi(X/s_X)$ where $\Phi$ is
the standard normal CDF\footnote{Hence, $\Phi_{s_X}$ is the CDF of a zero-mean
Gaussian with standard deviation $s_X$.}, $S(X)$ be the logistic function, and
$C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
distribution. Our goal is to solve:
\begin{equation}\label{eq:overall_corr_problem}
    C_\sigma^* = \argmin_{C_\sigma} |\Phi_{\sigma} * C_{\sigma} - S|
\end{equation}
where $*$ denotes convolution.

@@

With this approach, there is no guarantee that $u^* \geq 0$. However, we have
some flexibility in the choice of $\sigma$ in Equation~(\ref{eq:overall_corr_problem}).
As we decrease the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$
grows by the same amount and is in fact the result of convolution with a
Gaussian whose variance is the difference.  Thus as $\sigma$ decreases,
$C_\sigma(X)$ grows and approaches the derivative of a logistic function at
$\sigma = 0$. It retains some very weak negative values for $\sigma > 0$ but
removal of those values leads to very small error. Table~\ref{tab:xcorr} shows
that the errors between $X_{\rm norm}+X_{\rm corr}$ and $X_{\rm log}$ can be
made very small, approaching single floating precision (about $10^{-7}$), and
Algorithm~\ref{alg:our_algorithm} describes our procedure. A few points:

@@

We run MCMC with our M-H test and benchmark with~\citet{cutting_mh_2014}
and~\citet{icml2014c1_bardenet14}, all with initial minibatch size $m=100$.  For
the former, we increment minibatches by 100 and set the tolerance
$\epsilon=0.005$. For the latter, we increase sizes geometrically with a ratio
of $\gamma = 1.5$ and use error parameters $p = 2$ and $\delta = 0.01$.  All
methods collect 5000 samples using the same random walk proposer with covariance
${\rm diag}(0.15, 0.15)$, which means all shaping of distribution of the samples
is due to the M-H test.

@@

We first observe that
\[
    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
\]
and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
\begin{align*}
-\epsilon &= \int_{-\infty}^{+\infty} -\epsilon R(x) dx \\
&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon,
\end{align*}
as desired.
